String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-20 2023-1-19/sections/implementation.tex
remove some contents
crv version finished

%!TEX root = ../main.tex
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Does the decision procedure perform well~(Sec \ref{subsec:overall_eval}), especially when the counting bounds are large?~(Sec \ref{subsec:large_bounds_eval})
  \item [Q2:] Are the size-reduction techniques practical?~(Sec \ref{subsec:size_reduction_eval})
  \item [Q3:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker~\cite{atva2020}?~(Sec \ref{subsec:size_reduction_eval})
\end{itemize}

\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}

@@

\medskip
\noindent

\emph{Large-Bounds part.} This benchmark suite is composed of 1,969 problem instances with large counting bounds (greater than or equal to $50$) from RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.

@@

\subsection{Overall Evaluation}\label{subsec:overall_eval}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, and the results distinct to the ground truth are called \emph{soundness errors}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is \textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.

%
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the size-reduction techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \caption{Overall evaluation and the evaluation of the size-reduction techniques}
\end{table}

\subsection{Evaluation on problem instances with large bounds}\label{subsec:large_bounds_eval}
%
We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \caption{Experiment results on Large-Bounds part, with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\end{figure}

We evaluate the performance of $\ostrichrecl$ on 1,969 instances of the Large-Bounds part. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. To answer \textbf{Q1}, we can see that $\ostrichrecl$ \textbf{is far more efficient than the other solvers on problem instances with large bounds.}

\subsection{Evaluation of the effectiveness of size-reduction techniques}\label{subsec:size_reduction_eval}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$. To answer \textbf{Q2}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$. To answer \textbf{Q3}, \textbf{our algorithm for solving the nonemptiness problem is much better than the \textsc{nuXmv}-based techniques}.

>>>>

%!TEX root = ../main.tex

We implemented the algorithm in Section~\ref{sec:algorithm} on top of OSTRICH, resulted to a string solver called $\ostrichrecl$. 
%
In this section, we evaluate the performance of $\ostrichrecl$ on two benchmark suites, that is, RegCoL and AutomatArk. In the sequel, we first describe the two benchmark suites as well as the experiment setup. Then we present the experiment results. 
We do experiments to compare the performance of $\ostrichrecl$ with the state-of-the-art string solvers. Moreover, in order to know whether $\ostrichrecl$ is good at solving string constraints with large counting and length bounds, we extract 1,969 instances with large bounds out of the two benchmark suites, and compare the performance of $\ostrichrecl$ with the other solvers on these instances. Finally, we empirically justify the technical choices made in the decision procedure of Section~\ref{subsec:cefadec} by comparing $\ostrichrecl$ with the following two variants of $\ostrichrecl$: $\ostrichrecl_{\rm - ASR}$ and $\ostrichrecl_{\rm NUXMV}$, where $\ostrichrecl_{\rm - ASR}$ and $\ostrichrecl_{\rm NUXMV}$ are obtained from $\ostrichrecl$ by removing the automata size-reduction technique (i.e. Step 2 in Section~\ref{subsec:cefadec}) and using the nuXmv model checker to solve the nonemptiness of $\cefadec$ respectively.

%including the results for an overall evaluation as well as the results on the benchmark instances with large bounds. Moreover, we do experiments to show the effectiveness of the automata size-reduction technique (i.e. Step 2 in Section~\ref{sec:algorithm}) as well as compare $\ostrichrecl$ with the  nuXmv-based approach in \cite{atva2020}.

%Then we present experiment results of the large-counting part and evaluate the simplification techniques. The goal of our experiments is to answer the following questions:
%\begin{itemize}
%  \item [Q1:] Does the decision procedure perform well~(Sec \ref{subsec:overall_eval}), especially when the counting bounds are large?~(Sec \ref{subsec:large_bounds_eval})
%  \item [Q2:] Are the size-reduction techniques practical?~(Sec \ref{subsec:size_reduction_eval})
%  \item [Q3:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker~\cite{atva2020}?~(Sec \ref{subsec:size_reduction_eval})
%\end{itemize}

\subsection{Benchmark suites and experiment setup}\label{sec:bench}

@@

@@

%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=0.9\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
      \import{tables}{table_regcol.tex}
      \caption{Overall performance evaluation}
    \label{fig:table_overall_eval}
  \end{subfigure}
  \caption{Distribution of counting bounds and performance evaluation}
\end{figure}

\subsection{Performance evaluation}\label{subsec:overall_eval}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{Z3-str3}, Z3str3RE \cite{BD+23}, and OSTRICH
\cite{CHL+19}, on RegCoL and AutomatArk benchmark suites.
The experiment results can be found in Figure~\ref{fig:table_overall_eval}. Note that we take the results of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, and the results different from the ground truth are classified as \emph{soundness error}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves 48,662 instances correctly. The number is 3,908/1,111/12,838/2,306/2,396 more than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (1.93s versus 1.62s).

%To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.

%

\subsection{Evaluation on problem instances with large bounds}\label{subsec:large_bounds_eval}

We extract 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.

We evaluate the performance of $\ostrichrecl$ on the 1,969 instances. 
The experiment results can be found in Table~\ref{tab:large-bound}. We can see that $\ostrichrecl$ solves 1,873 instances correctly, which is 947/278/563/637/523 more than those solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Furthermore, $\ostrichrecl$ is 6.79/2.88/2.61/5.27/3.95 times faster than CVC5/Z3str3RE/ Z3str3/Z3seq/OSTRICH respectively. From the results, we can conclude that $\ostrichrecl$ is much more effective and efficient to solve the problem instances with large bounds than the other solvers.  
%To answer \textbf{Q1}, we can see that $\ostrichrecl$ \textbf{is far more efficient than the other solvers on problem instances with large bounds.}

\begin{table}
  \centering
  \begin{subtable}{0.53\textwidth}
      \centering
    \import{tables}{table_large_count.tex}
      \caption{Large bounds}
      \label{tab:large-bound}
  \end{subtable}
  \begin{subtable}{0.45\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Empirical justification of the technical choices in the decision procedure}
      \label{tab:results_simp}
  \end{subtable}
  \caption{More experiment results, where the time limit is set as 60 seconds}
\end{table}

\subsection{Empirical justification of the technical choices made in the decision procedure}\label{subsec:size_reduction_eval}
%
%We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
%Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
We compare $\ostrichrecl$ with $\ostrichrecl_{\rm -ASR}$ and $\ostrichrecl_{\rm NUXMV}$, to justify the technical choices made in Section~\ref{subsec:cefadec}. 
The experiment results can be found in Table~\ref{tab:results_simp}. 
We can see that $\ostrichrecl$ solves 1,503 more instances and is 2.21 times faster than $\ostrichrecl_{\rm -ASR}$. Therefore, the automata size-reduction technique indeed plays an essential role in the performance improvement. 
%To answer \textbf{Q2}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}. 
Moreover, $\ostrichrecl$ solves 1,798 more instances and is 3.13 times faster than $\ostrichrecl_{\rm NUXMV}$. Therefore, the decision procedure in Section~\ref{subsec:cefadec} is more efficient to solve the $\cefadec$ problem than nuXmv.

%To answer \textbf{Q3}, \textbf{our algorithm for solving the nonemptiness problem is much better than the \textsc{nuXmv}-based techniques}.

### 2023-10-20 2023-1-19/sections/introduction.tex
crv version finished

%!TEX root = ../main.tex
The string data type plays a crucial role in modern programming languages such as JavaScript, Python, Java, and PHP. 
String manipulations are error-prone and could even give rise to severe security vulnerabilities (e.g., cross-site scripting, aka XSS). 
One powerful method for identifying such bugs is \emph{symbolic execution}, which is possibly in combination with dynamic analysis. It analyses symbolic paths in a program by viewing them as constraints checked by constraint solvers. 
%
Symbolic execution of string manipulating programs has motivated the highly active research area of \emph{string constraint solving}, resulting in the development of numerous string solvers in the last decade, e.g.,
Z3seq~\cite{z3seq}, CVC4/5~\cite{cvc4}, Z3str/2/3/4~\cite{Z3-str,Z3-str2,Z3-str3,BerzishMurphy2021}, Z3str3RE~\cite{BD+23}, 
Z3-Trau~\cite{Z3-trau}\cite{z3trau}, OSTRICH~\cite{CHL+19}, Slent~\cite{WC+18}, among many others.

>>>>

%!TEX root = ../main.tex
The string data type plays a crucial role in modern programming languages such as JavaScript, Python, Java, and PHP. 
String manipulations are error-prone and could even give rise to severe security vulnerabilities (e.g., cross-site scripting, aka XSS). 
One powerful method for identifying such bugs is \emph{symbolic execution}, which is possibly in combination with dynamic analysis. It analyses symbolic paths in a program by viewing them as constraints checked by constraint solvers. 
%
Symbolic execution of string manipulating programs has motivated the highly active research area of \emph{string constraint solving}, resulting in the development of numerous string solvers in the last decade, e.g.,
Z3seq~\cite{z3seq}, CVC4/5~\cite{cvc4,cvc5}, Z3str/2/3/4~\cite{Z3-str,Z3-str2,Z3-str3,BerzishMurphy2021}, Z3str3RE~\cite{BD+23}, 
Z3-Trau~\cite{Z3-trau}\cite{z3trau}, OSTRICH~\cite{CHL+19}, Slent~\cite{WC+18}, among many others.
